Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-07-31 Paper/method.tex
-m 'changes'

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-based approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\tau$. 
\Ensure $\Lambda$, $W$, $\eta$. 
    \Procedure{Gradient-based}{}
	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$. 
    \State Set $N_\text{total}$ = $n_1$ 
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
	\State For each $k=1, \ldots, N_\text{total}$, compute $G(\bm\theta_k)$ and the gradient $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$
    \iffalse  
	\Statex\hspace{5mm} Using Finite Difference: 
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
         compute $\bm{g}^k$.
         \fi 
	\State Compute $\mat{C}$ and its eigenvalue decomposition 
		$\mat{C}$= $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top = W^{(0)}\Lambda^{(0)} W^{(0)\top}$
	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
	\State Partition the eigenpairs: $\Lambda^{(0)}~=~ 
        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
        $W^{(0)}~=~\begin{bmatrix} W_1^{(0)} & W_2^{(0)} \end{bmatrix}$, 
        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ \hayleynote{What is S?}

	\State Set $r$ = 0
	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. (\hayleynote{I am unsure what $\beta$ is? Maybe it has been defined somewhere else?})

	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
		\iffalse
		\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
		\State Compute $\mat{C}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
		\fi
		\State Repeat steps 4-7 above
		\State Compute $\delta W_{1,j}^{(r)}$ = 
                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{r})^2 - 
                       (W_{1,j}^{r-1})^2\|}{\|(W_{1,j}^{r-1})^2\|}$}, 
                       $j = 1,\ldots,\mathcal{S}$.
		\If {$\max\left(\delta W_{1,j}^{(r)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}w_{1_{i,j}}^2$,
	$i=1,\ldots,d$.

@@

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-free approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
\Ensure $\Lambda$, $W$, $\eta$. 
    \Procedure{Gradient-free}{}
    \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
    according to $\bm{f(\xi)}$.
    \State Set $r$ = 0
	\State Draw $n_r$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$. Set $N_\text{total}$ = $n_r$ 
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
	\State Construct the matrix $\mat{C}$.
	\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
	\State Partition the Eigenpairs: $\Lambda^{(r)}~=~ 
        \begin{bmatrix} \Lambda_1^{(r)} & \\ & \Lambda_2^{(r)} \end{bmatrix}$, 
        $W^{(r)}~=~\begin{bmatrix} W_1^{(r)} & W_2^{(r)} \end{bmatrix}$, 
        $\Lambda_1^{(r)}\in \mathbb{R}^{d\times\mathcal{S}}$
	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r$ \iffalse = $\lceil \beta n_1 \rceil$ \fi new random samples 
                $\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1] \iffalse $\beta = 2,\ldots $ $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi
        \State $N_\text{total}$ = $N_\text{total}$ + $n_r$   
        \State Repeat steps 5-9  
		\iffalse \State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.

@@

\bigskip
\begin{breakablealgorithm}
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{Algorithm for constructing the matrix, $\mat{C}$}
  \begin{algorithmic}[1]
	\Procedure{Local Linear Approximation}{}
	\State Draw $N$ random samples, $\{\bm{\xi}_k\}_{k=1}^{N}$ $\in$ [-1,1]
	according to $\bm{f(\xi)}$.
	\State Project to the physical space:
	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$
	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N$.
	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
	according to $\bm{f(\xi)}$.
	\State Choose an integer $p \leq N$ 
	\State For each $i=1, \ldots, M$, compute 
	\[
	\begin{aligned}
	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_k\}_{k=1}^{N} \text{ to } \bm{\nu}_i\}\\
	\vspace{-2mm}
	\Psi_i &= \text{subset of } G(\bm\theta_k) \text{ corresponding to the points in } \Phi_i
	\end{aligned}
	\]
	\State Compute linear least square fits 
	 $G(\bm\theta_k) \approx c_i + b_i^T\bm{\xi}_k$, \  $\bm{\xi}_k \in \Phi_i, \ G(\bm\theta_k) \in \Psi_i$
	\State Compute the matrix 
	\[
	\mat{C} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = {W}{\Lambda}{W}^T
	\]
	\EndProcedure
  \end{algorithmic}
  \label{alg:free2}
\end{breakablealgorithm}
\bigskip

>>>>

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-based approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\tau$. 
\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. %
    \Procedure{Gradient-based}{}
    \State Set $r$ = 0
	\State Draw $n_r$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$. 
    \State Set $N_\text{total}$ = $n_r$ 
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
	\State For each $k=1, \ldots, N_\text{total}$, compute $G(\bm\theta_k)$ and the gradient $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$
    \iffalse  
	\Statex\hspace{5mm} Using Finite Difference: 
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
         compute $\bm{g}^k$.
         \fi 
	\State Compute $\mat{C}$ and its eigenvalue decomposition 
		$\mat{C}$= $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top = W^{(r)}\Lambda^{(r)} W^{(r)\top}$
	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
	\State Partition the eigenpairs: $\Lambda^{(r)}~=~ 
        \begin{bmatrix} \Lambda_1^{(r)} & \\ & \Lambda_2^{(r)} \end{bmatrix}$, 
        $W^{(r)}~=~\begin{bmatrix} W_1^{(r)} & W_2^{(r)} \end{bmatrix}$, 
        $\Lambda_1^{(r)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ \hayleynote{What is S?}

	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r =  \beta n_{r-1}$  new random samples 
                $\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1], $\beta =??$ \iffalse $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi

	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
		\iffalse
		\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
		\State Compute $\mat{C}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
		\fi
		\State Repeat steps 5-8 above
		\State Compute $\delta W_{1,j}^{(r)}$ = 
                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{r})^2 - 
                       (W_{1,j}^{r-1})^2\|}{\|(W_{1,j}^{r-1})^2\|}$}, 
                       $j = 1,\ldots,\mathcal{S}$.
		\If {$\max\left(\delta W_{1,j}^{(r)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}W_{1_{i,j}}^2$,
	$i=1,\ldots,d$.

@@

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-free approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. 
    \Procedure{Gradient-free}{}
    \State Set $r$ = 0
	\State Draw $n_r$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$. Set $N_\text{total}$ = $n_r$ 
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
     \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
        according to $\bm{f(\xi)}$.
	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
	\State Construct the matrix $\mat{C}$.
	\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
	\State Partition the Eigenpairs: $\Lambda^{(r)}~=~ 
        \begin{bmatrix} \Lambda_1^{(r)} & \\ & \Lambda_2^{(r)} \end{bmatrix}$, 
        $W^{(r)}~=~\begin{bmatrix} W_1^{(r)} & W_2^{(r)} \end{bmatrix}$, 
        $\Lambda_1^{(r)}\in \mathbb{R}^{d\times\mathcal{S}}$
	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r =  \beta n_{r-1}$  new random samples 
		$\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1], $\beta =??$ \iffalse $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi
        \State $N_\text{total}$ = $N_\text{total}$ + $n_r$   
        \State Repeat steps 4-9  
		\iffalse \State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.

@@

\bigskip
\begin{breakablealgorithm}
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{Algorithm for constructing the matrix, $\mat{C}$}
  \begin{algorithmic}[1]
	\Procedure{Local Linear Approximation}{} \hayleynote{We could get rid of this algorithm completely by putting it in with algorithm 2?}
	\State Draw $N$ random samples, $\{\bm{\xi}_k\}_{k=1}^{N}$ $\in$ [-1,1]
	according to $\bm{f(\xi)}$.
	\State Project to the physical space:
	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$
	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N$.
	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
	according to $\bm{f(\xi)}$.
	\State Choose an integer $p \leq N$ 
	\State For each $i=1, \ldots, M$, compute 
	\[
	\begin{aligned}
	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_k\}_{k=1}^{N} \text{ to } \bm{\nu}_i\}\\
	\vspace{-2mm}
	\Psi_i &= \text{subset of } G(\bm\theta_k) \text{ corresponding to the points in } \Phi_i
	\end{aligned}
	\]
	\State Compute linear least square fits 
	 $G(\bm\theta_k) \approx c_i + b_i^T\bm{\xi}_k$, \  $\bm{\xi}_k \in \Phi_i, \ G(\bm\theta_k) \in \Psi_i$
	\State Compute the matrix 
	\[
	\mat{C} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = {W}{\Lambda}{W}^T
	\]
	\EndProcedure
  \end{algorithmic}
  \label{alg:free2}
\end{breakablealgorithm}
\bigskip
